---
title: "Telecom Customers Churn Analysis"
output: flexdashboard::flex_dashboard
runtime: shiny

---


```{r setup, include=FALSE}
library(flexdashboard)
library(shiny)
library(rmarkdown)
library(knitr)
library(Hmisc)
library(DT)
library(arsenal)
library(caret)
library(e1071)
library(data.table)
library(ggplot2)
library(kernlab)
library(corrplot)
library(cowplot)
library(car)
library(dplyr)
library(plyr)
library(ROSE)
library(ROCR)
library(gridExtra)
library(gridGraphics)
library(grid)

set.seed(1)
assignInNamespace("cedta.override", c(data.table:::cedta.override,"rmarkdown"), "data.table")


opts_chunk$set(echo = FALSE, comment="", warning = FALSE, message = FALSE, tidy.opts=list(width.cutoff=55), tidy = TRUE)

```


```{r read_data, include=FALSE}
#setwd("C:/Users/xujin/Desktop/5243 Data Science/final")
#churn <- fread(input = "Churn.csv", verbose = FALSE)
churn <- fread(input = "WA_Fn-UseC_-Telco-Customer-Churn.csv", verbose = FALSE)
# change the class of columns from char to factor
changeCols <- colnames(churn)[which(as.vector(churn[,lapply(.SD, class)]) == "character")]
churn[,(changeCols):= lapply(.SD, as.factor), .SDcols = changeCols]
# convert SeniorCitizen as a binary class with two levels 'yes' and 'no'
churn[,SeniorCitizen:=factor(SeniorCitizen, levels=c(0,1), labels=c("No", "Yes"))]
# delete missing records
# omit rows where 'x' has a missing value
churn <- na.omit(churn, cols="TotalCharges")
# remove CustomerId featuren column from the dataset
churn <- churn[,2:21]
```


```{r constants}
id.name <- "id"
gender.name <- "gender"
SeniorCitizen.name <- "SeniorCitizen"
Partner.name <- "Partner"
tenure.name <- "tenure"
PhoneService.name <- "PhoneService"
MultipleLines.name <- "MultipleLines"
InternetService.name <- "InternetService"
OnlineSecurity.name <- "OnlineSecurity"
OnlineBackup.name <- "OnlineBackup"
DeviceProtection.name <- "DeviceProtection"
TechSupport.name <- "TechSupport"
StreamingTV.name <- "StreamingTV"
StreamingMovies.name <- "StreamingMovies"
Contract.name <- "Contract"
PaperlessBilling.name <- "PaperlessBilling"
PaymentMethod.name <- "PaymentMethod"
MonthlyCharges.name <- "MonthlyCharges"
TotalCharges.name <- "TotalCharges"
Churn.name <- "Churn"

col.names <- names(churn)
plot.type <- c("fill","dodge","stack")
plot.type.name <- c("proportion","count","stack-count")

gender.choice <- unique(churn$gender)
seniorcitizen.choice <- unique(churn$SeniorCitizen)
partner.choice <- unique(churn$Partner)
Dependents.choice <- unique(churn$Dependents)
MultipleLines.choice <- unique(churn$MultipleLines)
InternetService.choice <- unique(churn$InternetService)
OnlineSecurity.choice <- unique(churn$OnlineSecurity)
OnlineBackup.choice <- unique(churn$OnlineBackup)
DeviceProtection.choice <- unique(churn$DeviceProtection)
TechSupport.choice <- unique(churn$TechSupport)
StreamingTV.choice <- unique(churn$StreamingTV)
StreamingMovies.choice <- unique(churn$StreamingMovies)
Contract.choice <- unique(churn$Contract)
PaperlessBilling.choice <- unique(churn$PaperlessBilling)
PaymentMethod.choice <- unique(churn$PaymentMethod)
tenure.max <- max(churn$tenure)
tenure.min <- min(churn$tenure)


```



```{r functions}
percentage.table <- function(x, digits){
  tab <- table(x)
  percentage.tab <- 100*tab/(sum(tab))
  rounded.tab <- round(x = percentage.tab, digits = digits)
  return(rounded.tab)
}

round.numerics <- function(x, digits){
  if(is.numeric(x)){
    x <- round(x = x, digits = digits)
  }
  return(x)
}
mean.diff <- function(x, y){
  return(mean(x, na.rm=TRUE) - mean(y, na.rm=TRUE))
}

```


Introduction
=====================================  


**Churn** happens when a customer stop doing business with a company. Telecommunications industry is one of the industries that particularly interested in the churn rate since customers usually have a relatively long contract with the companies and the customers have multiple choices. 

The raw data contains **`r churn[, .N]`** rows (customers) and **`r dim(churn)[2]`** columns (features). The “Churn” column is our target.  Most of our data are categorical like Churn, Gender, Dependents, Multiple Lines, Internet Service, Contract, Payment Method, etc. And the numerical data are like Tenure, Monthly Charges, Total Charges.

Five Tabs are provided in this dashboard:  
- **Introduction**: brief overview and functional descriptions   
- **Distribution**: display distributions for each feature in the dataset; barchart for categorical variables, and histogram for continuous variables  
- **Correlation**: display relationship between two features; barchart for two categorical variables, boxplot for categorical vs. continuous variable, and scatterplot for two continuous variables  
- **Churn**: compare the levels for each feature based on target "Churn"  
- **Prediction**: inputs the customer's information to help you predict whether the customer will churn or not  

Click on the tabs to see different reports.


Distribution
===================================

Row {data-height=500}
-------------------------------------

```{r dist_input}

sidebarLayout(
  sidebarPanel(
    width = 3,
    # selection list for single or multiple items selection
    selectInput(inputId="dist_var", label = "Select the Variable:", choices = col.names[1:20], selected = col.names[6]),
     # checkbox for percentage display
    checkboxInput(inputId = "show_percentages", label = "Show Percentages", value = TRUE),
     # numeric input control for decimal points
     numericInput(inputId = "percentage_digits", label = "Percentage Digits(0-6)", value = 1, min = 0, max = 6, step =1)
    
   ),
  
  mainPanel(
    width= 9,
     #caption
      h2(textOutput(outputId="distCaption")),
     #barchart
      plotOutput(outputId="distPlot")
    
  )
)


```


```{r dist_output}

#caption output
output$distCaption <- renderText({
  paste0("Distribution of ",input$dist_var)
  })

# chart output
output$distPlot<-renderPlot({
   # get the class for selected features
  var_class <- churn[,class(get(input$dist_var))]
  if (var_class=="factor"){
    
   tab <- percentage.table(x = churn[, get(input$dist_var)], input$percentage_digits)
   table<-as.data.table(tab)
    # barplot
   p<-ggplot(table,aes(x=unlist(table[, 1]),y=unlist(table[,2])))+ geom_bar(stat="identity",color="black",fill="lightblue")+ 
     labs(title= eval(input$dist_var), x="",y="percentage(%)") 
   
   # checkbox for percentage display
    if(input$show_percentages == TRUE){
      p+geom_text(aes(label=paste0(unlist(table[,2]),"%"),y=unlist(table[,2])+2.0), size=4)
    }else{
      p
    }
  }
  
  
  else{
     q<-ggplot(churn, aes(x=churn[,get(input$dist_var)])) + geom_histogram(aes(y=(..count..)/churn[,.N]*100),bins=10,position='dodge',color="black",fill="lightblue") +labs(title= eval(input$dist_var), x="", y="percentage(%)") 
    # checkbox for percentage display
    if(input$show_percentages == TRUE){
      q+stat_bin(aes(y=(..count..)/churn[,.N]*100, label=paste0(round((..count..)/churn[,.N]*100,input$percentage_digits),"%")), geom="text", size=4 ,bins=10, vjust=-0.75)
    }else{
      q
    }
   
  }
  
 
 })



```


Correlation
===================================


Row {data-height=500}
-------------------------------------

```{r correlation_input}

sidebarLayout(
  sidebarPanel(
    width = 3,
    # selection list for single or multiple items selection
    selectInput(inputId="variable1", label = "Select 1st Variable:", choices = col.names, selected = col.names[6]),
    selectInput(inputId="variable2", label = "Select 2nd Variable:", choices = col.names, selected = col.names[9])),
   
  
  mainPanel(
    width= 9,
     #caption
      h2(textOutput(outputId="corrCaption")),
      span(textOutput("helpInfo"), style="color:gray"),
     #barchart
      plotOutput(outputId="corrPlot")
    
  )
)


```


```{r correlation_output}

#caption output
output$corrCaption <- renderText({
  paste0(input$variable2, " vs. ",input$variable1)
  })

# chart output
output$corrPlot<-renderPlot({
  # get the class for selected features
  first_var_class <- churn[,class(get(input$variable1))]
  second_var_class <- churn[,class(get(input$variable2))]
  ## factor vs factor: bar chart
  if (first_var_class=='factor' & second_var_class=='factor'){
    #ggplot(churn, aes(churn[,get(input$variable1)],fill=get(input$variable2)))+geom_bar(position='fill') + labs(fill=eval(input$variable2), title= eval(input$variable1)) +xlab("") 
    
    ggplot(churn, aes(x= churn[,get(input$variable1)],  group=get(input$variable2))) + 
    geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") +
    geom_text(aes( label = scales::percent(..prop..),
                   y= ..prop.. ), stat= "count", vjust = -.5) +
    labs(x = eval(input$variable1),y="percentage",fill=eval(input$variable2), title=eval(input$variable2) ) +
  scale_y_continuous(labels = scales::percent)+facet_grid(~get(input$variable2))+
  theme(legend.title = element_blank(),legend.position = "none",axis.text.x = element_text(angle=30,hjust=1)) 
    
    
  }
  ## numeric vs factor: histogram
  else if(first_var_class!='factor' & second_var_class=='factor'){
     #ggplot(churn, aes(churn[,get(input$variable1)],fill=churn[,get(input$variable2)]))+geom_histogram(alpha=0.8,bins=20,position='dodge') + labs(fill=eval(input$variable2), title = eval(input$variable1), x="") 
    ggplot(data=churn, aes( x=churn[,get(input$variable2)], y = churn[,get(input$variable1)], fill = get(input$variable2))) + geom_boxplot() + labs(fill=eval(input$variable2), y=eval(input$variable1), x=eval(input$variable2)) + theme(axis.text.x=element_text(angle=30,hjust=1))
  }
  ## factor vs numeric : histogram
  else if (first_var_class=='factor' & second_var_class!='factor'){
    #ggplot(churn, aes(churn[,get(input$variable2)],fill=churn[,get(input$variable1)])) + geom_histogram(alpha=0.8,bins=20,position='dodge') + labs(fill=eval(input$variable1), title= eval(input$variable1), x="") 
    
    ggplot(data=churn, aes( x=churn[,get(input$variable1)], y = churn[,get(input$variable2)], fill = get(input$variable1))) + geom_boxplot() + labs(fill=eval(input$variable1), y=eval(input$variable2), x=eval(input$variable1)) + theme(axis.text.x=element_text(angle=30,hjust=1)) + coord_flip()
  }
  ## numeric vs numeric: scatter plot
  else{
    
     ggplot(data=churn, aes( x=churn[,get(input$variable2)], y = churn[,get(input$variable1)], color=2)) + geom_point() + labs(color="",y=eval(input$variable1), x=eval(input$variable2)) + theme(legend.position="none")
  }
  

})
#caption output
output$helpInfo <- renderText({
  # get the class for selected features
  first_var_class <- churn[,class(get(input$variable1))]
  second_var_class <- churn[,class(get(input$variable2))]
   ## factor vs factor: bar chart
  if (first_var_class=='factor' & second_var_class=='factor'){
    paste0("Use Barchart for categorical vs categorical.")
  }else if(first_var_class!='factor' & second_var_class=='factor'){
    paste0("Use Boxplot for numerical vs categorical.")
  }else if (first_var_class=='factor' & second_var_class!='factor'){
   paste0("Use Boxplot for categorical vs. numerical.")
  }else{
  paste0("Use ScatterPlot for numerical vs. numerical.")
  }
  
 })

```




Churn
===================================

Row {data-height=500}
-------------------------------------

```{r churn_input}

sidebarLayout(
  sidebarPanel(
    width = 3,
    # selection list for single or multiple items selection
    selectInput(inputId="churn_var", label = "Select the Variable:", choices = col.names[1:19], selected = col.names[6]),
    selectInput(inputId="plot_type", label = "Select the plot type:", choices = plot.type.name, selected = plot.type[2])
   ),
  
  mainPanel(
    width= 9,
     #caption
      h2(textOutput(outputId="churnCaption")),
     #barchart
      plotOutput(outputId="churnPlot"),
     #caption
      h2(textOutput(outputId="churnCaption2")),
     #summary table
      dataTableOutput(outputId="churnTable")
  )
  
)


```


```{r churn_output}

#caption output
output$churnCaption <- renderText({
  paste0("Churn vs. ",input$churn_var)
  })

# chart output
output$churnPlot<-renderPlot({
   # get the class for selected features
  var_class <- churn[,class(get(input$churn_var))]
  if (var_class=="factor"){
     ggplot(churn, aes(churn[,get(input$churn_var)],fill=get(Churn.name)))+geom_bar(position=plot.type[which(plot.type.name==input$plot_type)]) + labs(title= eval(input$churn_var), x=eval(input$churn_var),fill=eval(Churn.name),y="") 
    #ggplot(churn, aes(churn[,get(input$churn_var)]))+geom_bar(position=input$plot_type) + labs(title= eval(input$churn_var), x="") + facet_grid(cols = vars(Churn))
  }
  ## numeric: histogram
  else {
     ggplot(churn, aes(churn[,get(input$churn_var)],fill=get(Churn.name))) + geom_histogram(alpha=0.8,bins=20,position=plot.type[which(plot.type.name==input$plot_type)]) + labs(title= eval(input$churn_var), x=eval(input$churn_var), fill=eval(Churn.name),y="") 
  }
 })
 
  #caption output
output$churnCaption2 <- renderText({
  paste0("Statistics of ",input$churn_var, " group by Churn")
  }) 
 
 

output$churnTable<-DT::renderDataTable({
 
  # initial data summary
  #tab <- as.data.frame(summary(tableby(churn[,get(input$churn_var)]~churn[,Churn])))
  #tab <- as.data.table(tab)
  
   # get the class for selected features
  var_class <- churn[,class(get(input$churn_var))]
  if (var_class=="factor"){
    
     tab <- as.data.frame(summary(tableby(churn[,get(input$churn_var)]~churn[,Churn])))
     tab <- as.data.table(tab)
     level_name <- churn[,levels(get(input$churn_var))]
     #tab[,1:=c(input$churn_var,level_name)]
     tab[,1:=c("Churn","No","Yes")]
     # Rename a column in R
    colnames(tab)[1] <- eval(input$churn_var)
  }else{
     tab <- as.data.frame(summary(tableby(churn[,Churn]~churn[,get(input$churn_var)])))
     tab <- as.data.table(tab)
     tab[,1:=c(input$churn_var,"Mean(SD)","Range")]
     # Rename a column in R
    colnames(tab)[1] <- eval(Churn.name)
   
  }
  
 
  
  datatable(tab,rownames = FALSE)
  
})

```


Prediction
===================================

Row {data-height=500}
-------------------------------------

```{r train and test}
churn_copy=copy(churn)
#change variables to factors
col <-colnames(churn_copy)
factor.col <-col [churn_copy[,lapply(X=.SD,FUN="class")]=='factor']

indexes <- sample(1:nrow(churn_copy), size=0.3*nrow(churn_copy))

# drop TotalCharges column
churn_copy[,PhoneService:=NULL]
# drop TotalCharges column
churn_copy[,TotalCharges:=NULL]
train <- churn_copy[-indexes,]
```

```{r svm, include=FALSE}
train_control<- trainControl(method="cv", number=5, savePredictions = TRUE)

# train the model 
svm.mod.radial.cv <- train(Churn~., data=train, trControl=train_control, method="svmRadial")
```


```{r xgboost_input}
inputPanel(
  selectInput(inputId = "gender", label = "Enter gender:", 
              choices = gender.choice, selected = gender.choice[1]),
  selectInput(inputId = "senior_citizen", label = "Whether is senior citizen:", 
              choices = seniorcitizen.choice, selected = seniorcitizen.choice[1]),
  selectInput(inputId = "partner", label = "Enter partner:", 
              choices = partner.choice, selected = partner.choice[1]),
  selectInput(inputId = "dependents", label = "Whether has dependents:", 
              choices = Dependents.choice, selected = Dependents.choice[1]),
  sliderInput(inputId = "tenure", label = "Enter tenure length", min = tenure.min, max = tenure.max, value = 1, step =
1),
  selectInput(inputId = "MultipleLines", label = "Whether has multiple lines:", choices = MultipleLines.choice, selected = MultipleLines.choice[1]),
  selectInput(inputId = "InternetService", label = "select internet service:", 
              choices = InternetService.choice, selected = InternetService.choice[1]),
  selectInput(inputId = "OnlineSecurity", label = "If you have internet service, select online security", 
              choices = OnlineSecurity.choice, selected = OnlineSecurity.choice[1]),
  selectInput(inputId = "OnlineBackup", label = "If you have internet service, select online backup:", 
              choices = OnlineBackup.choice, selected = OnlineBackup.choice[1]),
  selectInput(inputId = "DeviceProtection", label = "Enter deviceProtection:", 
              choices = DeviceProtection.choice, selected = DeviceProtection.choice[1]),
  selectInput(inputId = "TechSupport", label = "If you have internet service, select tech support:", 
              choices = TechSupport.choice, selected = TechSupport.choice[1]),
  selectInput(inputId = "StreamingTV", label = "If you have internet service, select streamingTV:", 
              choices = StreamingTV.choice, selected = StreamingTV.choice[1]),
  selectInput(inputId = "StreamingMovies", label = "If you have internet service, select streamingMovies:", 
              choices = StreamingMovies.choice, selected = StreamingMovies.choice[1]),
  selectInput(inputId = "Contract", label = "select Contract:", 
              choices = Contract.choice, selected = Contract.choice[1]),
  selectInput(inputId = "PaperlessBilling", label = "select Paperless Billing:", 
              choices = PaperlessBilling.choice, selected = PaperlessBilling.choice[1]),
  selectInput(inputId = "PaymentMethod", label = "select Payment Method:", 
              choices = PaymentMethod.choice, selected = PaymentMethod.choice[1]),
  textInput(inputId = "MonthlyCharges", label = "Enter Monthly Charges:",
              value = "")
#,
  #textInput(inputId = "TotalCharges", label = "Enter Total Charges:",value = "")
)
```

```{r xgboost_output}
renderDataTable({
  example <- data.frame(gender=input$gender, SeniorCitizen=input$senior_citizen, Dependents =(input$dependents), Partner = input$partner, tenure =as.numeric(input$tenure), MultipleLines =(input$MultipleLines), 
InternetService =(input$InternetService), OnlineSecurity =(input$OnlineSecurity),OnlineBackup =(input$OnlineBackup), DeviceProtection = input$DeviceProtection, TechSupport =(input$TechSupport), StreamingTV =(input$StreamingTV), StreamingMovies =(input$StreamingMovies), Contract =(input$Contract), PaperlessBilling =(input$PaperlessBilling), PaymentMethod =(input$PaymentMethod), MonthlyCharges =as.numeric(input$MonthlyCharges)#, #TotalCharges =as.numeric(input$TotalCharges)
)
  
  pred <- data.frame(predict_churn = predict(svm.mod.radial.cv, example))
  datatable(pred)
})
```

